final-project

### 2024-11-08 final-project.Rmd
table 2 and diagnostics

```{r}
library(dplyr)
library(stringr)
library(psych)
library(gmodels)
library(survival)
```

@@

    # assign N stage type for easier analysis
    n2015 = case_when(
      str_detect(n2015, "N0") ~ "N0",
      str_detect(n2015, "N1|N1N0S") ~ "N1",
      str_detect(n2015, "N2|N2N0S") ~ "N2",
      str_detect(n2015, "N3|N3N0S") ~ "N3",
      str_detect(n2015, "NX") ~ "NX",
      TRUE ~ NA_character_
    ),
    n2017 = case_when(
      str_detect(n2017, "c0|p0") ~ "N0",
      str_detect(n2017, "c1|p1") ~ "N1",
      str_detect(n2017, "c2|p2") ~ "N2",
      str_detect(n2017, "c3|p3") ~ "N3",
      str_detect(n2017, "cX") ~ "NX",
      TRUE ~ NA_character_
    ),
    n2018 = case_when(
      str_detect(n2018, "N0") ~ "N0",
      str_detect(n2018, "N1|N1N0S") ~ "N1",
      str_detect(n2018, "N2|N2N0S") ~ "N2",
      str_detect(n2018, "N3|N3N0S") ~ "N3",
      str_detect(n2018, "NX") ~ "NX",
      TRUE ~ NA_character_
    ),
    # then merge the 2015, 2017, and 2018 N stage columns
    n_stage = coalesce(n2015, n2017, n2018),

@@

## age
summary(survfit(Surv(time = survival_months, event = survival_status) ~ age), times = c(12, 60, 120))
```

>>>>

```{r}
library(dplyr)
library(stringr)
library(psych)
library(gmodels)
library(survival)
library(survminer)
```

@@

    # assign N stage type for easier analysis
    n2015 = case_when(
      str_detect(n2015, "N0") ~ "N0",
      str_detect(n2015, "N1|N1N0S") ~ "N1",
      str_detect(n2015, "N2|N2N0S") ~ "N2",
      str_detect(n2015, "N3|N3N0S") ~ "N3",
      str_detect(n2015, "NX") ~ NA_character_,
      TRUE ~ NA_character_
    ),
    n2017 = case_when(
      str_detect(n2017, "c0|p0") ~ "N0",
      str_detect(n2017, "c1|p1") ~ "N1",
      str_detect(n2017, "c2|p2") ~ "N2",
      str_detect(n2017, "c3|p3") ~ "N3",
      str_detect(n2017, "cX") ~ NA_character_,
      TRUE ~ NA_character_
    ),
    n2018 = case_when(
      str_detect(n2018, "N0") ~ "N0",
      str_detect(n2018, "N1|N1N0S") ~ "N1",
      str_detect(n2018, "N2|N2N0S") ~ "N2",
      str_detect(n2018, "N3|N3N0S") ~ "N3",
      str_detect(n2018, "NX") ~ NA_character_,
      TRUE ~ NA_character_
    ),
    # then merge the 2015, 2017, and 2018 N stage columns
    n_stage = coalesce(n2015, n2017, n2018),

@@

## age
summary(survfit(Surv(time = survival_months, event = survival_status) ~ age), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ age)

## race
summary(survfit(Surv(time = survival_months, event = survival_status) ~ race), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ race)

## marriage status
summary(survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ marriage_status)

## N stage
summary(survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ n_stage)

## M stage
summary(survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ m_stage)

## grade
summary(survfit(Surv(time = survival_months, event = survival_status) ~ grade), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ grade)

## radiation
summary(survfit(Surv(time = survival_months, event = survival_status) ~ radiation), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ radiation)

## chemotherapy
summary(survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ chemotherapy)

## surgery
summary(survfit(Surv(time = survival_months, event = survival_status) ~ surgery), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ surgery)
```

```{r}
# check the proportional hazards assumption for each covariate

## age
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ age), data = seer, fun="cloglog")

## race
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ race), data = seer, fun="cloglog")

## marriage status
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), data = seer, fun="cloglog")

## N stage
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), data = seer, fun="cloglog")

## M stage
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), data = seer, fun="cloglog")

## grade
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ grade), data = seer, fun="cloglog")

## radiation
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ radiation), data = seer, fun="cloglog")

## chemotherapy
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), data = seer, fun="cloglog")

## surgery
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ surgery), data = seer, fun="cloglog")
```
